Spatiotemporal Multi-Resolution Approximation of the Amari Type Neural Field Model

### 2011-11-14 Multiresolution_framework/ltx/manuscript.tex
minor correction

By differentiating the $\mathcal Q$-function with respect to $\boldsymbol\theta$ we have
\begin{align}\label{eq:MRA-QDerivative}
\frac{\partial \mathcal Q}{\partial \boldsymbol\theta}&=2T_s\varsigma(\boldsymbol\upsilon_0-\boldsymbol\upsilon_1)^\top-T_s\varsigma(\boldsymbol\Upsilon^\top+\boldsymbol\Upsilon)\boldsymbol\theta \nonumber \\
&=2T_s\varsigma(\boldsymbol\upsilon_0-\boldsymbol\upsilon_1)^\top-2T_s\varsigma\boldsymbol\Upsilon^\top\boldsymbol\theta.
\end{align}
Equating \eqref{eq:MRA-QDerivative} to $\mathbf 0$ yields
\begin{align}\label{eq:MRA-thetahat}
\boldsymbol \theta= \boldsymbol\Upsilon^{-\top}(\boldsymbol\upsilon_0-\boldsymbol\upsilon_1)^\top.
\end{align}
The second derivative of the $\mathcal Q$-function is given by
\begin{equation}
\frac{\partial^2\mathcal Q}{\partial\boldsymbol\theta^2}=-2T_s\varsigma\boldsymbol\Upsilon.
\end{equation}
The matrix $\boldsymbol\Upsilon$ is positive definite if $n_x^2\times n_{\theta}$ matrix $\mathrm{vec}(\mathbf U)$ is of rank $n_{\theta}$ (\cite{Dewar2009}, Lemma 2), where $\mathrm{vec}(\cdot)$ denotes the vectorisation operator. Under this condition $\boldsymbol\Upsilon$ is  invertible and the second derivative is negative definite representing a maximum of the $\mathcal Q$-function.

@@

For each of the experiment, the estimation was applied to the final 900~ms, allowing the model's dynamics to stabilise from the initial conditions. The initial state, parameters and the support of the connectivity kernel were unknown to the estimator. The observation noise was set to $\boldsymbol\Sigma_{\varepsilon}=0.1 \times \mathbf{I}_{n_y}$ and the disturbance covariance function was set to $\gamma(r-r') = \varphi_{3,-2}(r-r')$ (B-spline, $j=3$).  The firing rate slope, $\varsigma$ was set to $0.56~mV^{-1}$ \cite{Wendling2005}. The observation kernel was modelled using a B-spline function, with a width of 0.8~mm at half the maximum amplitude where the distance between adjacent sensors was $0.5$~mm, resulting into $n_y = 161$ observations. The spacing and bandwidth of the sensors allowed the full spatial bandwidth of the field to be observed.  With this number of observations, the estimation problem was well-posed, i.e. the number of states, $n_x < n_y$ at each level of the decomposition. This is because the scaling functions are orthogonal to the wavelets at the same and higher levels of decomposition, and  the wavelets are orthogonal to each other across different scales (see (\ref{eq:MRA-PsiPsiOrthogonality}-\ref{eq:MRA-PhiPsiOrthogonality})). 
\subsection{Experiment I}
The spatial frequency response of the observed field was used to specify the level of decomposition required to represent the field using the wavelets. This is shown in Figure~\ref{fig:ObservationFrequencyResponce} where the graph obtained by averaging the spectral power of the spatial frequency (over time) from the observations \cite{Scerri2009}. The wavelets at each spatial scale was expanded in terms of the scaling function at the next scale using \eqref{eq:MRA-TwoScalepair2} and the resulting expansion was then used to analytically calculate the spatial bandwidth  via \eqref{eq:MRA-NmFouriertransform}. The result suggested that wavelets up to level $j=3$ (with the bandwidth $\approx[5,8]$ cycles/mm) can represent the significant spatial characteristics of the field which yields $n_x = 131$ states, corresponding to 9 scaling functions at $j=0$ and 8, 16, 32 and 66 wavelet functions respectively at $j=0$, $j=1$, $j=2$ and $j=3$. 
\begin{figure}[!h] 
 	\centering
 		\includegraphics[scale=1]{./Graph/Figure2.pdf}
 		\caption{{\bf Spatial frequency analysis}. The average (over time) power in dB of the spatial frequency of the observations.}
 	\label{fig:ObservationFrequencyResponce}
 \end{figure}

>>>>

By differentiating the $\mathcal Q$-function with respect to $\boldsymbol\theta$ we have
\begin{align}\label{eq:MRA-QDerivative}
\frac{\partial \mathcal Q}{\partial \boldsymbol\theta}&=2T_s\varsigma(\boldsymbol\upsilon_0-\boldsymbol\upsilon_1)^\top-T_s\varsigma(\boldsymbol\Upsilon^\top+\boldsymbol\Upsilon)\boldsymbol\theta \nonumber \\
&=2T_s\varsigma(\boldsymbol\upsilon_0-\boldsymbol\upsilon_1)^\top-2T_s\varsigma\boldsymbol\Upsilon^\top\boldsymbol\theta.
\end{align}
Equating \eqref{eq:MRA-QDerivative} to $\mathbf 0$ yields
\begin{align}\label{eq:MRA-thetahat}
\boldsymbol \theta= \boldsymbol\Upsilon^{-\top}(\boldsymbol\upsilon_0-\boldsymbol\upsilon_1)^\top.
\end{align}
The second derivative of the $\mathcal Q$-function is given by
\begin{equation}
\frac{\partial^2\mathcal Q}{\partial\boldsymbol\theta^2}=-2T_s\varsigma\boldsymbol\Upsilon.
\end{equation}
The matrix $\boldsymbol\Upsilon$ is positive definite if $n_x^2\times n_{\theta}$ matrix $\mathrm{vec}(\mathbf U)$ is of rank $n_{\theta}$ (\cite{Dewar2009}, Lemma 3), where $\mathrm{vec}(\cdot)$ denotes the vectorisation operator. Under this condition $\boldsymbol\Upsilon$ is  invertible and the second derivative is negative definite representing a maximum of the $\mathcal Q$-function.

@@

For each of the experiment, the estimation was applied to the final 900~ms, allowing the model's dynamics to stabilise from the initial conditions. The initial state, parameters and the support of the connectivity kernel were unknown to the estimator. The observation noise was set to $\boldsymbol\Sigma_{\varepsilon}=0.1 \times \mathbf{I}_{n_y}$ and the disturbance covariance function was set to $\gamma(r-r') = \varphi_{3,-2}(r-r')$ (B-spline, $j=3$).  The firing rate slope, $\varsigma$ was set to $0.56~mV^{-1}$ \cite{Wendling2005}. The observation kernel was modelled using a B-spline function, with a width of 0.8~mm at half the maximum amplitude where the distance between adjacent sensors was $0.5$~mm, resulting into $n_y = 161$ observations. The spacing and bandwidth of the sensors allowed the full spatial bandwidth of the field to be observed.  With this number of observations, the estimation problem was well-posed, i.e. the number of states, $n_x < n_y$ at each level of the decomposition. This is because the scaling functions are orthogonal to the wavelets at the same and higher levels of decomposition, and  the wavelets are orthogonal to each other across different scales (see (\ref{eq:MRA-PsiPsiOrthogonality}-\ref{eq:MRA-PhiPsiOrthogonality})). 
\subsection{Experiment I}
The spatial frequency response of the observed field was used to specify the level of decomposition required to represent the field using the wavelets. This is shown in Figure~\ref{fig:ObservationFrequencyResponce} where the graph obtained by averaging the spectral power of the spatial frequency (over time) from the observations \cite{Scerri2009}. The spatial bandwidths of wavelets was calculated analytically using \eqref{eq:MRA-NmFouriertransform}. The result suggested that wavelets up to level $j=3$ (with the bandwidth $\approx[5,8]$ cycles/mm) can represent the significant spatial characteristics of the field which yields $n_x = 131$ states, corresponding to 9 scaling functions at $j=0$ and 8, 16, 32 and 66 wavelet functions respectively at $j=0$, $j=1$, $j=2$ and $j=3$. 
\begin{figure}[!h] 
 	\centering
 		\includegraphics[scale=1]{./Graph/Figure2.pdf}
 		\caption{{\bf Spatial frequency analysis}. The average (over time) power in dB of the spatial frequency of the observations.}
 	\label{fig:ObservationFrequencyResponce}
 \end{figure}
